### 2024-07-18 evaluating-prediction-accuracy.qmd
20240718 - miscalibration

@@

If you focus on only one accuracy index, [RMSE](#sec-rootMeanSquaredError) would be a good choice.
However, I would also examine a [calibration plot](#sec-calibrationPlot) to evaluate whether predictions are poorly calibrated at higher levels of points.
I would also examine [ME](#sec-meanError)—not to compare the accuracy of various predictions per se—but to determine whether predictions are systematically under- or overestimating actual points.
If so, predictions may be able to be refined by adding or subtracting a constant to the predictions; however, this could worsen other accuracy indices, so it is important to conduct an interative process of modifying then evaluating, then further modifiying and evaluating, etc.
It may also be valuable to evaluate the accuracy of various subsets of the predictions.
For instance, you might examine the predictive accuracy of players whose project points are greater than 100, to evaluate the accuracy of predictions specifically to distinguish between players at higher levels of points.

@@

- [discrimination](#sec-discrimination):
  - [area under the receiver operating curve](#sec-auc)
  - and, secondarily—depending on the particulary cutoff and the relative costs of false positives versus false negatives:
    - [sensitivity](#sec-sensitivity)
    - [specificicity](#sec-specificity)
    - [positive predictive value](#sec-ppv)
    - [negative predictive value](#sec-npv)
- [calibration](#sec-calibration):
  - [calibration plot](#sec-calibrationPlot)
  - [Spiegelhalter's *z*](#sec-spiegelhalterZ)

>>>>

#### {#sec-miscalibration}

The mean difference between predicted values versus observed values at a given cutoff is an index of [miscalibration](#sec-calibration) of predictions at that cutoff.
It is called "calibration-in-the-small" (as opposed to calibration-in-the-large, which spans all cutoffs).
Values closer to zero reflect greater accuracy.
Values above zero indicate that the predicted values are, on average, greater than the observed values.
Values below zero indicate that the observed values are, on average, greater than the predicted values.

@@

If you focus on only one accuracy index, [RMSE](#sec-rootMeanSquaredError) would be a good choice.
However, I would also examine a [calibration plot](#sec-calibrationPlot) to evaluate whether predictions are poorly calibrated at higher levels of points.
I would also examine [ME](#sec-meanError)—not to compare the accuracy of various predictions per se—but to determine whether predictions are systematically under- or overestimating actual points.
If so, predictions may be able to be refined by adding or subtracting a constant to the predictions; however, this could worsen other accuracy indices, so it is important to conduct an iterative process of modifying then evaluating, then further modifiying and evaluating, etc.
It may also be valuable to evaluate the accuracy of various subsets of the predictions.
For instance, you might examine the predictive accuracy of players whose project points are greater than 100, to evaluate the accuracy of predictions specifically to distinguish between players at higher levels of points.

@@

- [discrimination](#sec-discrimination):
  - [area under the receiver operating curve](#sec-auc)
  - and, secondarily—depending on the particulary cutoff and the relative costs of false positives versus false negatives:
    - [sensitivity](#sec-sensitivity)
    - [specificicity](#sec-specificity)
    - [positive predictive value](#sec-ppv)
    - [negative predictive value](#sec-npv)
- [calibration](#sec-calibration):
  - [calibration plot](#sec-calibrationPlot)
  - [Spiegelhalter's *z*](#sec-spiegelhalterZ)
  - [Mean difference between observed and predicted values](#sec-miscalibration)
